```{r}
usage_data = read.csv('BatMobile_Usage_20703267.csv')
length(usage_data$B)
```

## question 1a

check for null values

```{r}

sum(is.na(usage_data))

```

get averages of each feature and standard deviation

```{r}
print('A -')
print(sd(usage_data$A))
summary(usage_data$A)
print('B -')
print(sd(usage_data$B))
summary(usage_data$B)
print('C -')
print(sd(usage_data$C))
summary(usage_data$C)

boxplot(usage_data)


```

plot box plots These box plots show that B has an outlier which is pulling up the mean. On first inspection it would seem that group b is the strongest however first we must consider this outlier.

after removing the outlier and second outlier has been observed in feature B. As this outlier is only just below the lower quartile I've decided n

```{r}
usage_data$B[which(usage_data$B == 690)] = median(usage_data$B)

getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

print('A -')
print(sd(usage_data$A))
print(getmode(usage_data$A))
summary(usage_data$A)
print('B -')
print(sd(usage_data$B))
print(getmode(usage_data$B))
summary(usage_data$B)
print('C -')
print(sd(usage_data$C))
print(getmode(usage_data$C))
summary(usage_data$C)

boxplot(usage_data)

```

scatter graphs

plot(usage_data$A)
plot(usage_data$B, add.scope())

## Question 1b

distribution

```{r}



#hist(usage_data$B, freq=FALSE)
df = length(usage_data$B)-1

curve(dt(x,df=df),-3,3,ylab="density",col='purple') # plot a T_{n-1} density
abline(v=qt(0.85,df=df)) # add a vertical line at the 0.975 quantile (i.e. the probability of exceeding is 2.5%)
abline(v=qt(0.15,df=df))
print(qt(0.925,df=df))
print(qt(0.075,df=df))


```

```{r}
test = t.test(usage_data$B, conf.level = 0.85, mu=mean(usage_data$B))
test
```

## Question 1C

```{r}
hist(usage_data$A, freq=FALSE, main='Histogram of manufacturer A Gamma dist')
curve(dgamma(x,21,0.0523), 200, 700, col='purple', add=TRUE)

```

```{r}

hist(usage_data$A, freq=FALSE, main='Histogram of manufacturer A Normal dist')
curve(dnorm(x,mean(usage_data$A),sd(usage_data$A)), 200, 700, col='red', add=TRUE)
```

```{r}

#calculate observed intervals for the data
cat<-seq(200,600,by=50)
observed<-seq(200,600,by=50)
for(i in 2:9){
  observed[i-1]<-sum(usage_data$A>cat[i-1]&usage_data$A<cat[i])
} 
observed[9]<-sum(usage_data$A>600&usage_data$A<650) # the last category 


#calculate expected outcomes 
prob<-seq(200,600,by=50) # A vector to store the probability of the first goal being in each of the 7 intervals for the data
for(i in 2:9){
  prob[i-1]=pgamma(cat[i],21,0.0523)-pgamma(cat[i-1],21,0.0523)
}
prob[9]<-pgamma(650,21,0.0523)-pgamma(cat[9],21,0.0523)
expected<-round(prob*55)
  

print(expected)
print(observed)

plot(cat,observed, xlab="Groupings of batter failers (minuets)",ylab="Frequency",ylim=c(0,14))
points(cat,expected, col=2, pch=20)
legend(500,10,legend=c("Observed","Expected"),col=c(1,2),pch=c(20,20))
```

```{r}
chi_test=sum((observed-expected)^2/expected) 


df = 9-1-2 # -2 because we used an estimated alpha and beta
round(qchisq(0.90,df),3)
round(qchisq(0.95,df),3)
chi_test
```

## question 2a

```{r}
comparison_data = read.csv('BatMobile_Comparison_20703267.csv')
```

```{r}
head(comparison_data)
```

```{r}
print(sum(is.na(comparison_data)))
print(is.na(comparison_data))
```

```{r}

boxplot(comparison_data[2:3])
```

```{r}
max(comparison_data$Day)
min(comparison_data$Day)
length(comparison_data$Day)
```

```{r}
plot(comparison_data$Day, comparison_data$TurboBM, xlab="Day",ylab="Minuets")
points(comparison_data$Day, comparison_data$BMClassic, col=2, pch=20)
for (i in comparison_data$Day){
  abline(v=i, col='Purple')
}
legend(30,650,legend=c("TurboBM","BMClassic"),col=c(1,2),pch=c(1,20))
```

```{r}

print('A -')
print(sd(na.omit(comparison_data$TurboBM)))
print(getmode(na.omit(comparison_data$TurboBM)))
summary(comparison_data$TurboBM)
print('B -')
print(sd(na.omit(comparison_data$BMClassic)))
print(getmode(na.omit(comparison_data$BMClassic)))
summary(comparison_data$BMClassic)

```

```{r}
alpha = round(mean(na.omit(comparison_data$TurboBM))^2/var(na.omit(comparison_data$TurboBM)),4)
beta = round(mean(na.omit(comparison_data$TurboBM))/var(na.omit(comparison_data$TurboBM)),4)
hist(comparison_data$TurboBM, freq=FALSE)
curve(dgamma(x,alpha,beta), 200, 700, col='red', add=TRUE)
```

```{r}
alpha = round(mean(na.omit(comparison_data$BMClassic))^2/var(na.omit(comparison_data$BMClassic)),4)
beta = round(mean(na.omit(comparison_data$BMClassic))/var(na.omit(comparison_data$BMClassic)),4)
hist(comparison_data$BMClassic, freq=FALSE)
curve(dgamma(x,alpha,beta), 150, 700, col='red', add=TRUE)
```

## Question 2b

```{r}
cor.test(comparison_data$TurboBM,comparison_data$BMClassic)
```

```{r}
correlation<-seq(1,5, by=1)
library(dplyr)
for (i in 1:5){
  sampled_data<-sample_n(na.omit(comparison_data[2:3]),10)
  print(sampled_data)
  correlation[i] <- cor(sampled_data$TurboBM, sampled_data$BMClassic)
}
correlation
```

## questopm 2c

```{r}
var.test(comparison_data$TurboBM,comparison_data$BMClassic)
```

## question 2d

```{r}
t.test(comparison_data$TurboBM,comparison_data$BMClassic, paired=TRUE, mu=55) 
t.test(comparison_data$TurboBM,comparison_data$BMClassic, paired=TRUE, mu=55, conf.level=0.90) 
```

## question 3a

```{r}
lab_data = read.csv('BatMobile_Laboratory_20703267.csv')
head(lab_data)
print(sum(is.na(lab_data)))
```

```{r}
model1=lm(lab_data$Lifetime~lab_data$Temperature)
lifetime_temp<-data.frame(lab_data$Temperature,lab_data$Lifetime)
plot(lifetime_temp, col='red',pch=20)
abline(a=model1$coef[1],b=model1$coef[2])
```

```{r}

range(lifetime_temp$lab_data.Temperature)

```

```{r}
mean_temp<-seq(1,5)

mean_temp[1] = mean(lifetime_temp$lab_data.Lifetime[lifetime_temp$lab_data.Temperature < 0])
mean_temp[2] = mean(lifetime_temp$lab_data.Lifetime[lifetime_temp$lab_data.Temperature < 10 & lifetime_temp$lab_data.Temperature > 0])
mean_temp[3] = mean(lifetime_temp$lab_data.Lifetime[lifetime_temp$lab_data.Temperature < 20 & lifetime_temp$lab_data.Temperature > 10])
mean_temp[4] = mean(lifetime_temp$lab_data.Lifetime[lifetime_temp$lab_data.Temperature < 30 & lifetime_temp$lab_data.Temperature > 20])
mean_temp[5] = mean(lifetime_temp$lab_data.Lifetime[lifetime_temp$lab_data.Temperature > 30])

mean_temp
```

```{r}
lab_data$Temperature = ((lab_data$Temperature-15)^2)
model2=lm(lab_data$Lifetime~lab_data$Temperature)
lifetime_temp<-data.frame(lab_data$Temperature,lab_data$Lifetime)
plot(lifetime_temp, col='red',pch=20)
abline(a=model2$coef[1],b=model2$coef[2])
```

```{r}
summary(model2)
```

## Question 3b

```{r}
pair_plot<-lab_data[2:6]
print(pair_plot)
pairs(data.frame(lab_data$Temperature,lab_data$Recharge,lab_data$Memory, lab_data$Lifetime))
plot(lab_data$Lifetime, lab_data$Memory, col='red',pch=20)
```

```{r}
model3<-lm(Lifetime~., data=lab_data)
summary(model3) 
AIC(model3)
```

```{r}
hist(model3$residuals,xlab="Residuals",main="",freq=FALSE)
curve(dnorm(x,0,sd(model3$residuals)),-40,80,col=2,add=TRUE) 
plot(model3$fitted.values,model3$residuals,xlab="Fitted values",ylab="Residuals",col='red',pch=20)
abline(h=0,lty=2)
mean(model3$residuals)
median(model3$residuals)
getmode(model3$residuals)
```

```{r}
mean((model3$residuals))
```



## best model bits

```{r}
bestmodel3<-step(model3) # Try removing each covariate in turn.  Stop when no further reduction in AIC can be found
summary(bestmodel3) # Display a summary of the model with the lowest AIC
AIC(bestmodel3)
print(bestmodel3$coefficients)
```

```{r}
hist(bestmodel3$residuals,xlab="Residuals",main="",freq=FALSE)
curve(dnorm(x,0,sd(bestmodel3$residuals)),-50,80,col=2,add=TRUE) 
plot(bestmodel3$fitted.values,bestmodel3$residuals,xlab="Fitted values",ylab="Residuals",col='red',pch=20)
print(bestmodel3$coefficients)
abline(h=0,lty=2)
```

```{r}
model_recharge=lm(lab_data$Lifetime~lab_data$Recharge)
lifetime_recharge<-data.frame(lab_data$Recharge,lab_data$Lifetime)
plot(lifetime_recharge, col='red',pch=20)
abline(a=model_recharge$coef[1],b=model2$coef[2])
```

```{r}
boxplot(lab_data$Lifetime~lab_data$Device)
boxplot(lab_data$Lifetime~lab_data$Level)
```

```{r}
model_test<-lm(Lifetime~Temperature+Device+Level, data=lab_data)
summary(model_test) 
AIC(model_test)
```


```{r}
anova_model <- lm(Lifetime~Device, data=lab_data)
anova(bestmodel3)
```

```{r}
pred_cost=predict(bestmodel3,newdata=data.frame(Recharge=37, Device='B', Level='Low', Temperature=(23.7-15)^2),interval='prediction', level=0.75)
pred_cost
```




